### 2021-05-13 manuscript/article.tex

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

\title{Evolution, functional diversity, regime shifts, hysteresis, and
nonlinearity; ecosystem responses to environmental change.}

\author{
    Owen L Petchey
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:owen.petchey@ieu.uzh.ch}{\nolinkurl{owen.petchey@ieu.uzh.ch}}} \\
   \And
    Rainer Krug
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:rainer.krug@ieu.uzh.ch}{\nolinkurl{rainer.krug@ieu.uzh.ch}}} \\
   \And
    Marcel Suleiman
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:marcel.suleiman@ieu.uzh.ch}{\nolinkurl{marcel.suleiman@ieu.uzh.ch}}} \\
  }

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
To be completed
\end{abstract}

\keywords{
    Biodiversity
   \and
    Evolution
   \and
    Hysteresis
   \and
    Alternate stable states
   \and
    Environmental change
   \and
    Regime shift
   \and
    Threshold
   \and
    Ecosystem
   \and
    Cyanobacteria
   \and
    Sulfur bacteria
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

When thinking about ecosystem stability, a key question is how
biodiversity influences ecosystem responses to environmental change
(figure 1: panel a shows linear response with different gradients, large
being with low diversity; panel b shows a linear and a non-linear
response starting and ending at the same place; panel c shows a
hysteretic response, with greater or less range of hysteresis).

In this context, a directly relevant aspect of biodiversity is variation
in functional traits, i.e.~the traits that determine organismal
responses to environmental conditions and that determine the effects of
organisms on their environment (citation). Variation in functional
traits is often termed ``functional diversity.''

There is a considerable amount of theory predicting, and evidence
supporting, that greater functional diversity leads to lower temporal
variability in ecosystem states variables (such as total biomass)
(citations). The mechanism by which this occurs is that greater
functional response diversity causes a weaker relationship between an
environmental driver gradient and the ecosystem state (figure 1a), due
to the functional turnover along the environmental driver gradient that
results from variation in organisms' responses to the environmental
driver gradient .

There is less theory about effects of functional diversity on the
linearity (figure 1b) and the hysteresis (figure 1c) of ecosystem
responses to environmental change. Dakos et al. (2019), in their article
``Ecosystem tipping points in an evolving world'' asked how changes in
functional trait distributions (including evolutionary changes) affect
the occurrence of tipping points. They predict that:

\begin{itemize}
\item
  In general, greater functional response diversity may decrease the
  likelihood of a catastrophic ecosystem response (Figure 2a in Dakos et
  al. (2019)). Here, we interpret this prediction to be that greater
  functional response diversity decreases the non-linearity of ecosystem
  response to environmental driver change, including the extreme
  non-linearity that can occur when an ecosystem ``flips'' from one
  state to another one.
\item
  Greater functional trait diversity can make a tipping point occur only
  with greater change in an environmental driver (Figure 2a in Dakos et
  al. (2019)). I.e. functional trait diversity can increase
  resilience--the extent of environmental change required to push the
  ecosystem into another state (Holling (1973)). This could occur if
  greater functional diversity equates to the presence of organisms with
  greater tolerance to the environmental driver Dakos et al. (2019).
\item
  On the other hand, the opposite could occur: greater functional
  diversity could make a tipping point occur with less change in an
  environmental driver (i.e.~may cause lower resilience) (Figure 2c in
  @dakos2019b). This could occur if there is selection for greater
  tolerance is causes, incidentally, selection for lower growth rate,
  perhaps due to a tradeoff between the two traits.
\item
  Finally, the amount of functional diversity may affect the ``recovery
  trajectory,'' i.e.~the relationship between ecosystem state and the
  environmental driver when the driver change is reversed Dakos et al.
  (2019). Put another way, functional diversity may increase or decrease
  the amount of hysteresis, and may even result in a recovery with lower
  non-linearity / no tipping-point (Figure 3 in Dakos et al. (2019)).
\end{itemize}

One reason why these predictions are often indeterminate is that trait
distributions can be affected by, and effect, organismal abundances and
environmental conditions. These numerous resulting feedbacks make
predictions from verbal argumentation rather difficult. \textbf{Discuss
Ceulemans, R., Wojcik, L.A. \& Gaedke, U. (2021). Functional diversity
alters the effects of a pulse perturbation on the dynamics of tritrophic
food webs. \emph{bioRxiv}, 2021.03.22.436420.}

What of the empirical evidence? Hillebrand\ldots{}

\hypertarget{scope-of-our-investigation}{%
\subsection{Scope of our
investigation}\label{scope-of-our-investigation}}

Questions: - How does variation organismal tolerances affect response

Approach: - Modelling the system with ODEs. - Diversity manipulation
(can be considered selection among strains, or selection among species).
- Do not include mutation.

According the the questions (in the Introduction) this is what we did:

Variation in amount of among strain variation, no variation in number of
strains, no variation in strength or shape of the of the tradeoff

Effect of among strain variation on how the steady state(s) of the
system varies with in different environmental conditions, and on the
dynamical response of system to gradual environmental change.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{the-study-ecosystem}{%
\subsection{The study ecosystem}\label{the-study-ecosystem}}

Anoxic-oxic ecosystem shifts as a case study. Contemporary relevance.
Intermediate complexity of system\ldots{} diagram of system, including
some panels of subsystems showing some of the positive feedbacks.
Central role of inhibition / tolerance. Also showing responses to
changes in oxygen diffusivity (justify from Bush et al this as a key
environmental driver).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{/Users/owenpetchey/Desktop/microxanox/diversity_envresp1/manuscript/ecosystem_network}

}

\caption{your caption}\label{fig:ecosys_network}
\end{figure}

The model ecosystem contains three types (functional groups) of organism
(Figure \ref(fig:ecosys\_network)): cyanobacteria (\(CB\)),
sulfate-reducing bacteria (\(SB\)), and phototrophic sulfur bacteria
(\(PB\)), four chemical substrates (oxygen \(O\), phosphorous \(P\),
reduced sulfur \(SR\), and oxidised sulfur \(SO\)), and four types of
flow/interaction (production, consumption, inhibition, and diffusion).
Particularly important among these interactions are that the
cyanobacteria produce oxygen which inhibits the growth of both types of
sulfur bacteria. And the sulfate-reducing bacteria produce reduced
sulfur, such as hydrogen sulphide, which inhibit the growth of
cyanobacteria. This mutual inhibition (a positive feedback) creates the
potential for an oxic cyanobacteria dominated state and an anoxic sulfur
bacteria dominated state to be alternate stable states.

Simulation of this ecosystem as a set of ordinary differential equations
demonstrated that the presence of alternate stable states depended on
the rate at which oxygen diffuses into (or out of) the system (Bush et
al. 2017). Low levels of oxygen diffusivity result in only the anoxic
state being stable. High levels result in only the oxic state.
Intermediate oxygen diffusivity allowed alternate stable states, with
the occurring state depending on historical conditions. Historical
dominance of cyanobacteria caused the oxic state to occur, which
historical dominance of the sulfur bacteria caused the anoxic state to
occur. All this was thoroughly and elegantly presented in Bush et al.
(2017).

As mentioned above, critical for the presence of alternate stable states
in this ecosystem is the mutual inhibition between cyanobacteria and
sulfur bacteria. This mutual inhibition is created by their intolerance
of the chemical substrates that the other produces. Recent experimental
evolution studies have demonstrated considerable increases in tolerance.
For example, the sulfur bacteria \emph{Desulfovibrio vulgaris} evolved a
32-fold increase in oxygen tolerance, via relatively few mutations
(Schoeffler et al. 2019). And the cyanobacterium \emph{Microcystis
aeruginosa} evolved a four-fold higher sulfide tolerance via rare
spontaneous mutations (Martín-Clemente et al. 2019). In that study,
populations with higher genetic variation and those experiencing slower
environmental change (an increase in sulfide concentration) were more
likely to persist at high sulfide concentration. Furthemore, there is
considerable interspecific variation in tolerances (Schoeffler et al.
2019)

Physiological tolerances to environmental conditions, including
interspecific variation in tolerance and evolution of tolerance, are a
key component of the engineering of ecosystem for determining their
stability (Minervini \emph{et al.} 2014)(Gómez‐Gras \emph{et al.}
2019)(Cuenca Cambronero \emph{et al.} 2018)(Vos \emph{et al.}
2017)(Evans \& Wallenstein 2014)(Rolfe \emph{et al.} 1978)(Knoll \&
Bauld 1989)(Girvan \emph{et al.} 2005).

(Rolfe et al. 1978) - Factors related to the oxygen tolerance of
anaerobic bacteria. (Ramel et al. 2015) - ``they were able to grow but
the final biomasses and the growth yield were lower than that obtained
under anaerobic conditions,'' ``Determination of the molar growth yields
on lactate suggested that a part of the energy gained from lactate
oxidation was derived toward cells protection/repairing against
oxidative conditions rather than biosynthesis'' (Hamilton et al. 2018) -
``Cyanobacterial photosynthesis under sulfidic conditions: insights from
the isolate Leptolyngbya sp. strain hensonii''

\hypertarget{the-model}{%
\subsection{The model}\label{the-model}}

Bush et al (2017) contains an accessible and complete description of the
model of the ecosystem, including rate equations, parameter values, and
initial conditions. These features are also represented and documented
in the \texttt{microxanox} R package (\textbf{cite the package}).
Therefore, here we only describe the relatively straightforward
extension we used to model multiple strains per functional group.

Each of the strains within each functional groups has its own parameter
set and state variable. Thus a system with nine strains within each
functional group has 27 state variables, plus another four (one for each
of the four chemical substrates). Other than having potentially
different parameter values (e.g.~maximum growth rate and tolerance
parameter) the dynamics of the strains in a functional were described by
the same equations. Hence, when parameterised with multiple strains per
functional group, but no variation among strains (within functional
groups), the system behaves identically to when there is only one strain
per functional group (see the \texttt{microxanox} R package \emph{User
Guide} for a demonstration).

The only other necessary modification was to include summation terms in
each of the four ordinary differential equations describing the rates of
change of the four chemical substrates. This was required in order to
appropriately account for the production and consumption of these
substrates by multiple strains.

\hypertarget{creating-within-functional-group-diversity}{%
\subsection{Creating within functional group
diversity}\label{creating-within-functional-group-diversity}}

In this section we describe how we created variability among the
multiple strains within functional groups. Since the process was the
same for each functional group, we usually do not in this description
subscript parameter names by functional group (though there were
functional group specific parameters in all simulations). The
descriptions in this section gradually become more mathematical; the
next paragraph alone should provide the information required to
understand all other elements of this report.

We introduced among-strain variability in maximum growth rate (parameter
\(G_{max}\)) and in the tolerance to the relevant inhibiting chemical
substrate (parameter \(h\)). Among-strain variability was created such
that there was a tradeoff between maximum growth rate and tolerance
(i.e.~high tolerance coinciding with low maximum growth rate)
(e.g.~Figure @ref(fig:tradeoff). Our implementation of the variation and
tradeoff contain the assumption of a linear tradeoff of log-log
transformed among-strain variation.

For further understanding, please consider a reference maximum growth
rate parameter \(G_{max}\) and reference tolerance parameter \(h\).
Furthermore, consider the meta-parameters \(Div_{G_{max}}\) and
\(Div_{h}\), which determine the amount of variation among \(n\)-strains
in the growth rate and tolerance parameters. To create a tradeoff, the
two meta-parameters were always of different sign.

The among strain variation was calculated such that the growth rate of
strain \(i = 1\) was a factor \(1/(2^{Div_{G_{max}}})\) of the reference
growth rate, and the growth rate of the \(i = n\) strain was a factor
\(2^{Div_{G_{max}}}\) of the reference growth rate. The growth rate of
the \(i = \{2,...n-1\}\) other strains was the reference growth rate
multiplied by a factor that was equally distributed between the factor
of the \(i = 1\) and \(i = n\) strain.

Hence, for example, with the reference growth rate \(G_{max} =\) 0.05,
among strain variation of \(Div_{G_{max}} =\) 1, and nine strains
(\(n =\) 9), then the growth rates of the nine strains are 0.025, 0.03,
0.035, 0.042, 0.05, 0.059, 0.071, 0.084, 0.1 (Figure @ref(fig:tradeoff).

(Put another way, the growth rate parameter of the i-th of the \(n\)
strains in a functional group was calculated as
\(G_{max, i} = D_i * G_{max}\) where D\_i is the i-th element of the set
\(D = \{2^{f(x)Div_{G_{max}}} | x \in \mathbb{N}, x \leq n\}\), where
\(f(x) = (x-(n+1)/2) / (( n-1)/2)\).)

\hypertarget{measurement-of-non-linearity-and-hysteresis}{%
\subsection{Measurement of non-linearity and
hysteresis}\label{measurement-of-non-linearity-and-hysteresis}}

Requires an environmental change\ldots{} as in Bush, oxygen diffusivity.

Take from Garnier et al.

Rate-independent hysteresis, because we are considering only stable
states, and not rate-dependent hysteresis, which can occur when a lag in
the system response to an environmental response causes a difference
between the observed system state at time \(t\) and the stable state for
the environmental conditions present at time \(t\).

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

We implemented the model in the \texttt{microxanox} package
(\textbf{cite the package}) of the R programming language (**cite R Core
Team\emph{). The package contains functions for initialising, running,
visualising, and analysing simulations of the system described above.
Please refer to the \texttt{microxanox} package} User Guide* and
\emph{Partial reproduction of Bush et al (2017)} vignettes for further
information about it.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{figure-stable-states-without-variation}{%
\subsection{Figure: Stable states without
variation}\label{figure-stable-states-without-variation}}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{article_files/figure-latex/ss_novar-1}

}

\caption{your caption}\label{fig:ss_novar}
\end{figure}

X-axis must be labelled log10(a).

\hypertarget{figure-stable-states-with-variation-and-tradeoff}{%
\subsection{Figure: Stable states with variation and
tradeoff}\label{figure-stable-states-with-variation-and-tradeoff}}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{article_files/figure-latex/ss_var-1}

}

\caption{your caption}\label{fig:ss_var}
\end{figure}

X-axis must be labelled log10(a).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{article_files/figure-latex/ss_var2-1}

}

\caption{your caption}\label{fig:ss_var2}
\end{figure}

X-axis must be labelled log10(a).

\hypertarget{figure-hysteresis-shift-point-nonlinearity-as-var-increases.}{%
\subsection{Figure: hysteresis, shift point, nonlinearity, as var
increases.}\label{figure-hysteresis-shift-point-nonlinearity-as-var-increases.}}

Resolution of a\_grid with 100 steps is about 0.03 units, which is about
the resolution of the y axis here. Hysteresis range is measures in
log10(a) units.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{article_files/figure-latex/ss_var3-1}

}

\caption{your caption}\label{fig:ss_var3}
\end{figure}

\hypertarget{figure-temporal-switches}{%
\subsection{Figure: Temporal switches}\label{figure-temporal-switches}}

To be added. Need to go both ways (oxic to anoxic, and anoxic to oxic).

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

What does this study tell us about effects of functional diversity on
ecosystem stability?

What does this study tell us about the effects of evolution on ecosystem
stability?

Need a more process based approach to variation in biodiversity:
mutation, genetics,

As mentioned in the Introduction, the numerous feedacks among trait
distributions, organismal abundances, and environmental conditions,
resulting feedbacks make predictions from verbal argumentation rather
difficult. They may also make for context dependent effects of
functional diversity. I.e. functional diversity may delay a tipping
point in one context (e.g.~ecosystem type) but delay it in another. This
is why we need more investigations, both theoretical and empirical, of
effects of functional trait diversity on environmental responses to
environmental change.

\url{https://onlinelibrary.wiley.com/doi/10.1111/gcb.15662?af=R}

\url{https://onlinelibrary.wiley.com/doi/10.1111/ele.13760?af=R}

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

SNF URPP

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-bush2017}{}%
Bush, Timothy, Muhe Diao, Rosalind J. Allen, Ruben Sinnige, Gerard
Muyzer, and Jef Huisman. 2017. {``Oxic-Anoxic Regime Shifts Mediated by
Feedbacks Between Biogeochemical Processes and Microbial Community
Dynamics.''} \emph{Nature Communications} 8 (1): 789.
\url{https://doi.org/10.1038/s41467-017-00912-x}.

\leavevmode\hypertarget{ref-dakos2019b}{}%
Dakos, Vasilis, Blake Matthews, Andrew P. Hendry, Jonathan Levine,
Nicolas Loeuille, Jon Norberg, Patrik Nosil, Marten Scheffer, and Luc De
Meester. 2019. {``Ecosystem Tipping Points in an Evolving World.''}
\emph{Nature Ecology \& Evolution} 3 (3): 355--62.
\url{https://doi.org/10.1038/s41559-019-0797-2}.

\leavevmode\hypertarget{ref-hamilton2018}{}%
Hamilton, Trinity L., Judith M. Klatt, Dirk de Beer, and Jennifer L.
Macalady. 2018. {``Cyanobacterial Photosynthesis Under Sulfidic
Conditions: Insights from the Isolate Leptolyngbya Sp. Strain
Hensonii.''} \emph{The ISME Journal} 12 (2): 568--84.
\url{https://doi.org/10.1038/ismej.2017.193}.

\leavevmode\hypertarget{ref-holling1973}{}%
Holling, C S. 1973. {``Resilience and Stability of Ecological
Systems.''} \emph{Annual Review of Ecology and Systematics} 4 (1):
1--23. \url{https://doi.org/10.1146/annurev.es.04.110173.000245}.

\leavevmode\hypertarget{ref-martuxednclemente2019}{}%
Martín-Clemente, Elena, Ignacio J. Melero-Jiménez, Elena Bañares-España,
Antonio Flores-Moya, and María J. García-Sánchez. 2019. {``Adaptation
Dynamics and Evolutionary Rescue Under Sulfide Selection in
Cyanobacteria: A Comparative Study Between {\emph{Microcystis
Aeruginosa}} and {\emph{Oscillatoria}} Sp. (Cyanobacteria).''}
\emph{Journal of Phycology} 55 (6): 1348--60.
\url{https://doi.org/10.1111/jpy.12911}.

\leavevmode\hypertarget{ref-ramel2015}{}%
Ramel, Fanny, Gael Brasseur, Laetitia Pieulle, Odile Valette, Agnès
Hirschler-Réa, Marie Laure Fardeau, and Alain Dolla. 2015. {``Growth of
the Obligate Anaerobe Desulfovibrio Vulgaris Hildenborough Under
Continuous Low Oxygen Concentration Sparging: Impact of the
Membrane-Bound Oxygen Reductases.''} \emph{PLOS ONE} 10 (4): e0123455.
\url{https://doi.org/10.1371/journal.pone.0123455}.

\leavevmode\hypertarget{ref-rolfe1978}{}%
Rolfe, Rial D., David J. Hentges, Benedict J. Campbell, and James T.
Barrett. 1978. {``Factors Related to the Oxygen Tolerance of Anaerobic
Bacteria.''} \emph{Applied and Environmental Microbiology} 36 (2):
306--13. \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC291219/}.

\leavevmode\hypertarget{ref-schoeffler2019}{}%
Schoeffler, Marine, Anne-Laure Gaudin, Fanny Ramel, Odile Valette, Yann
Denis, Wagdi Ben Hania, Agnès Hirschler-Réa, and Alain Dolla. 2019.
{``Growth of an Anaerobic Sulfate-Reducing Bacterium Sustained by Oxygen
Respiratory Energy Conservation After O {\textsubscript{2}} -Driven
Experimental Evolution: O {\textsubscript{2}} -Driven Experimental
Evolution of {\emph{Desulfovibrio}}.''} \emph{Environmental
Microbiology} 21 (1): 360--73.
\url{https://doi.org/10.1111/1462-2920.14466}.

\end{CSLReferences}

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}