### 2021-06-29 experiments/tradeoff strength illustration.Rmd

---
title: "Experiment 1"
author: "Owen Petchey"
date: "6/25/2021"
output: html_document
---

# Setup

## R

```{r}

rm(list = ls())

knitr::opts_knit$set(progress = TRUE, verbose = TRUE, cache = TRUE)

# devtools::install_github("opetchey/microxanox",
#                              ref="main",
#                              auth_token = "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",
#                              build_vignettes = FALSE,
#                              force = TRUE)

library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("experiments/r functions/various_useful_functions.r"))
zero <- 0 ## don't change
unity <- 1 ## don't change!!!
options(mc.cores = 8)

colfunc_CB <- colorRampPalette(c("#B5FFC9", "#024F17"))
colfunc_SB <- colorRampPalette(c("#FCBEB3", "#7D1402"))
colfunc_PB <- colorRampPalette(c("#F9AEFC", "#6E0172"))

```

## General simulation conditions

```{r}
default_dynamic_model <- bushplus_dynamic_model
default_event_definition <- event_definition_1
default_event_interval <- 100
default_noise_sigma <- 0
default_minimum_abundances <- rep(1, 3)
names(default_minimum_abundances) <- c("CB", "PB", "SB") 
default_sim_duration <- 80000
default_sim_sample_interval <- 100
initial_pars_from <- "bush_ssfig3"
## note that next line (log10a_series is over-ridden with getting stable states)
#default_log10a_series <- c(-2, -2, -2, -2, -10, -10, -10, -10, -10)
```

## Define diversity

```{r}
num_CB_strains <- 9
num_SB_strains <- 9
num_PB_strains <- 9
CB_var_range <- 0.015789474
CB_var_TO <- -20
SB_var_range <- 0.1
SB_var_TO <- -20
PB_var_range <- 0.1
PB_var_TO <- -20

num_div_treatment_levels <- 10
```

## Create diversity

```{r}
default_9strain <- new_starter(n_CB = num_CB_strains,
                             n_SB = num_SB_strains,
                             n_PB = num_PB_strains,
                             values_initial_state = initial_pars_from)
CB_var_gmax_s <- seq(zero, unity*CB_var_TO, length=num_div_treatment_levels) * CB_var_range
CB_var_h_s = seq(zero, unity, length=num_div_treatment_levels) * CB_var_range
SB_var_gmax_s <- seq(zero, unity, length=num_div_treatment_levels) * SB_var_range
SB_var_h_s = seq(zero, SB_var_TO, length=num_div_treatment_levels) * SB_var_range
PB_var_gmax_s <- seq(zero, unity, length=num_div_treatment_levels) * PB_var_range
PB_var_h_s = seq(zero, PB_var_TO, length=num_div_treatment_levels) * PB_var_range
var_expt <- tibble(CB_var_gmax_s,
                   CB_var_h_s,
                   SB_var_gmax_s,
                   SB_var_h_s,
                   PB_var_gmax_s,
                   PB_var_h_s)
var_expt <- var_expt %>%
  nest_by(CB_var_gmax_s, CB_var_h_s,
          SB_var_gmax_s, SB_var_h_s,
          PB_var_gmax_s, PB_var_h_s) %>%
  mutate(pars = list(add_strain_var(default_9strain,
                                    CB_var_gmax = CB_var_gmax_s,
                                    CB_var_h = CB_var_h_s,
                                    SB_var_gmax = SB_var_gmax_s,
                                    SB_var_h = SB_var_h_s,
                                    PB_var_gmax = PB_var_gmax_s,
                                    PB_var_h = PB_var_h_s)))
```

## Display diversity

```{r}

CBtraits <- tibble(var_expt$pars[[10]]$CB,
                     Diversity = "None",
                   Strength = "Strong")
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[6]]$CB,
                     Diversity = "Medium",
                     Strength = "Strong"))
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[1]]$CB,
                     Diversity = "High",
                     Strength = "Strong"))

````

## Define diversity

```{r}
num_CB_strains <- 9
num_SB_strains <- 9
num_PB_strains <- 9
CB_var_range <- 0.015789474
CB_var_TO <- -2
SB_var_range <- 0.1
SB_var_TO <- -20
PB_var_range <- 0.1
PB_var_TO <- -20

num_div_treatment_levels <- 10
```

## Create diversity

```{r}
default_9strain <- new_starter(n_CB = num_CB_strains,
                             n_SB = num_SB_strains,
                             n_PB = num_PB_strains,
                             values_initial_state = initial_pars_from)
CB_var_gmax_s <- seq(zero, unity*CB_var_TO, length=num_div_treatment_levels) * CB_var_range
CB_var_h_s = seq(zero, unity, length=num_div_treatment_levels) * CB_var_range
SB_var_gmax_s <- seq(zero, unity, length=num_div_treatment_levels) * SB_var_range
SB_var_h_s = seq(zero, SB_var_TO, length=num_div_treatment_levels) * SB_var_range
PB_var_gmax_s <- seq(zero, unity, length=num_div_treatment_levels) * PB_var_range
PB_var_h_s = seq(zero, PB_var_TO, length=num_div_treatment_levels) * PB_var_range
var_expt <- tibble(CB_var_gmax_s,
                   CB_var_h_s,
                   SB_var_gmax_s,
                   SB_var_h_s,
                   PB_var_gmax_s,
                   PB_var_h_s)
var_expt <- var_expt %>%
  nest_by(CB_var_gmax_s, CB_var_h_s,
          SB_var_gmax_s, SB_var_h_s,
          PB_var_gmax_s, PB_var_h_s) %>%
  mutate(pars = list(add_strain_var(default_9strain,
                                    CB_var_gmax = CB_var_gmax_s,
                                    CB_var_h = CB_var_h_s,
                                    SB_var_gmax = SB_var_gmax_s,
                                    SB_var_h = SB_var_h_s,
                                    PB_var_gmax = PB_var_gmax_s,
                                    PB_var_h = PB_var_h_s)))
```

```{r}
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[10]]$CB,
                     Diversity = "None",
                     Strength = "Weak"))
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[6]]$CB,
                     Diversity = "Medium",
                     Strength = "Weak"))
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[1]]$CB,
                     Diversity = "High",
                     Strength = "Weak"))

ggplot(CBtraits) +
  geom_point(aes(x = h_SR_CB, y = g_max_CB)) +
  facet_grid(Strength ~ Diversity) +
  xlab("Tolerance trait") +
  ylab("Maximum growth rate")

```