### 2021-06-30 experiments/tradeoff strength illustration.Rmd

---
title: "Experiment 1"
author: "Owen Petchey"
date: "6/25/2021"
output: html_document
---

# Setup

## R

```{r}

rm(list = ls())

knitr::opts_knit$set(progress = TRUE, verbose = TRUE, cache = TRUE)

# devtools::install_github("opetchey/microxanox",
#                              ref="main",
#                              auth_token = "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",
#                              build_vignettes = FALSE,
#                              force = TRUE)

library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("experiments/r functions/various_useful_functions.r"))
zero <- 0 ## don't change
unity <- 1 ## don't change!!!
options(mc.cores = 8)

colfunc_CB <- colorRampPalette(c("#B5FFC9", "#024F17"))
colfunc_SB <- colorRampPalette(c("#FCBEB3", "#7D1402"))
colfunc_PB <- colorRampPalette(c("#F9AEFC", "#6E0172"))

```

## General simulation conditions

```{r}
default_dynamic_model <- bushplus_dynamic_model
default_event_definition <- event_definition_1
default_event_interval <- 100
default_noise_sigma <- 0
default_minimum_abundances <- rep(1, 3)
names(default_minimum_abundances) <- c("CB", "PB", "SB") 
default_sim_duration <- 80000
default_sim_sample_interval <- 100
initial_pars_from <- "bush_ssfig3"
## note that next line (log10a_series is over-ridden with getting stable states)
#default_log10a_series <- c(-2, -2, -2, -2, -10, -10, -10, -10, -10)
```

## Define diversity

```{r}
num_CB_strains <- 9
num_SB_strains <- 9
num_PB_strains <- 9
CB_gmax_div <- 0.015789474
CB_h_div <- -0.08
SB_gmax_div <- 0.015789474
SB_h_div <- -0.323
PB_gmax_div <- 0.015789474
PB_h_div <- -0.323

CB_var_range <- 0.015789474
CB_var_TO <- -20
SB_var_range <- 0.1
SB_var_TO <- -20
PB_var_range <- 0.1
PB_var_TO <- -20

num_div_treatment_levels <- 10
```

## Create diversity

```{r}
default_9strain <- new_starter(n_CB = num_CB_strains,
                             n_SB = num_SB_strains,
                             n_PB = num_PB_strains,
                             values_initial_state = initial_pars_from)
CB_var_gmax_s <- seq(zero, unity, length=num_div_treatment_levels) * CB_gmax_div
CB_var_h_s = seq(zero, unity, length=num_div_treatment_levels) * CB_h_div
SB_var_gmax_s <- 1#seq(zero, unity, length=num_div_treatment_levels) * SB_gmax_div
SB_var_h_s = 1#seq(zero, unity, length=num_div_treatment_levels) * SB_h_div
PB_var_gmax_s <- 1#seq(zero, unity, length=num_div_treatment_levels) * PB_gmax_div
PB_var_h_s = 1#seq(zero, unity, length=num_div_treatment_levels) * PB_h_div
var_expt <- crossing(CB_var_gmax_s,
                   CB_var_h_s,
                   SB_var_gmax_s,
                   SB_var_h_s,
                   PB_var_gmax_s,
                   PB_var_h_s)
var_expt <- var_expt %>%
  nest_by(CB_var_gmax_s, CB_var_h_s,
          SB_var_gmax_s, SB_var_h_s,
          PB_var_gmax_s, PB_var_h_s) %>%
  mutate(pars = list(add_strain_var(default_9strain,
                                    CB_var_gmax = CB_var_gmax_s,
                                    CB_var_h = CB_var_h_s,
                                    SB_var_gmax = SB_var_gmax_s,
                                    SB_var_h = SB_var_h_s,
                                    PB_var_gmax = PB_var_gmax_s,
                                    PB_var_h = PB_var_h_s)))
```

## Display diversity

```{r}

med_gmax <- 0.007017544
med_h <- -0.035555556
high_gmax <- 0.015789474
high_h <- -0.080

CBtraits <- tibble(var_expt$pars[[which(var_expt$CB_var_gmax_s == 0 &
                                          var_expt$CB_var_h_s == 0)]]$CB,
                     Div_gmax = "None",
                   Div_h = "None")
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[which(var_expt$CB_var_gmax_s == 0 & abs(var_expt$CB_var_h_s - med_h)<0.0001)]]$CB,
                     Div_gmax = "None",
                   Div_h = "Medium"))
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[which(var_expt$CB_var_gmax_s == 0 & abs(var_expt$CB_var_h_s - high_h)<0.0001)]]$CB,
                     Div_gmax = "None",
                   Div_h = "High"))

CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[which(abs(var_expt$CB_var_gmax_s - med_gmax)<0.0001 & abs(var_expt$CB_var_h_s - 0)<0.0001)]]$CB,
                     Div_gmax = "Medium",
                   Div_h = "None"))
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[which(abs(var_expt$CB_var_gmax_s - med_gmax)<0.0001 & abs(var_expt$CB_var_h_s - med_h)<0.0001)]]$CB,
                     Div_gmax = "Medium",
                   Div_h = "Medium"))
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[which(abs(var_expt$CB_var_gmax_s - med_gmax)<0.0001 & abs(var_expt$CB_var_h_s - high_h)<0.0001)]]$CB,
                     Div_gmax = "Medium",
                   Div_h = "High"))

CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[which(abs(var_expt$CB_var_gmax_s - high_gmax)<0.0001 & abs(var_expt$CB_var_h_s - 0)<0.0001)]]$CB,
                     Div_gmax = "High",
                   Div_h = "None"))
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[which(abs(var_expt$CB_var_gmax_s - high_gmax)<0.0001 & abs(var_expt$CB_var_h_s - med_h)<0.0001)]]$CB,
                     Div_gmax = "High",
                   Div_h = "Medium"))
CBtraits <- bind_rows(CBtraits,
                      tibble(var_expt$pars[[which(abs(var_expt$CB_var_gmax_s - high_gmax)<0.0001 & abs(var_expt$CB_var_h_s - high_h)<0.0001)]]$CB,
                     Div_gmax = "High",
                   Div_h = "High"))

```

## Define diversity

```{r}
# num_CB_strains <- 9
# num_SB_strains <- 9
# num_PB_strains <- 9
# CB_var_range <- 0.015789474
# CB_var_TO <- -2
# SB_var_range <- 0.1
# SB_var_TO <- -20
# PB_var_range <- 0.1
# PB_var_TO <- -20
# 
# num_div_treatment_levels <- 10
```

## Create diversity

```{r}
# default_9strain <- new_starter(n_CB = num_CB_strains,
#                              n_SB = num_SB_strains,
#                              n_PB = num_PB_strains,
#                              values_initial_state = initial_pars_from)
# CB_var_gmax_s <- seq(zero, unity*CB_var_TO, length=num_div_treatment_levels) * CB_var_range
# CB_var_h_s = seq(zero, unity, length=num_div_treatment_levels) * CB_var_range
# SB_var_gmax_s <- seq(zero, unity, length=num_div_treatment_levels) * SB_var_range
# SB_var_h_s = seq(zero, SB_var_TO, length=num_div_treatment_levels) * SB_var_range
# PB_var_gmax_s <- seq(zero, unity, length=num_div_treatment_levels) * PB_var_range
# PB_var_h_s = seq(zero, PB_var_TO, length=num_div_treatment_levels) * PB_var_range
# var_expt <- tibble(CB_var_gmax_s,
#                    CB_var_h_s,
#                    SB_var_gmax_s,
#                    SB_var_h_s,
#                    PB_var_gmax_s,
#                    PB_var_h_s)
# var_expt <- var_expt %>%
#   nest_by(CB_var_gmax_s, CB_var_h_s,
#           SB_var_gmax_s, SB_var_h_s,
#           PB_var_gmax_s, PB_var_h_s) %>%
#   mutate(pars = list(add_strain_var(default_9strain,
#                                     CB_var_gmax = CB_var_gmax_s,
#                                     CB_var_h = CB_var_h_s,
#                                     SB_var_gmax = SB_var_gmax_s,
#                                     SB_var_h = SB_var_h_s,
#                                     PB_var_gmax = PB_var_gmax_s,
#                                     PB_var_h = PB_var_h_s)))
```

```{r}
# CBtraits <- bind_rows(CBtraits,
#                       tibble(var_expt$pars[[10]]$CB,
#                      Diversity = "None",
#                      Strength = "Weak"))
# CBtraits <- bind_rows(CBtraits,
#                       tibble(var_expt$pars[[6]]$CB,
#                      Diversity = "Medium",
#                      Strength = "Weak"))
# CBtraits <- bind_rows(CBtraits,
#                       tibble(var_expt$pars[[1]]$CB,
#                      Diversity = "High",
#                      Strength = "Weak"))

ggplot(CBtraits) +
  geom_point(aes(x = h_SR_CB, y = g_max_CB)) +
  facet_grid(Div_gmax ~ Div_h) +
  xlab("Tolerance trait") +
  ylab("Maximum growth rate")

```