### 2021-11-26 experiments/0_ss_finding/Error_Test.Rmd
Error isolation and solution steps

---
title: "Test Error"
author: "Rainer M Krug"
date: "11/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  error = TRUE,
  messages = TRUE
)

library(here)
source(here("R/various_useful_functions.r"))

num_strains <- 9
num_CB_strains <- num_strains
num_SB_strains <- num_strains
num_PB_strains <- num_strains

source(here("experiments/0_ss_finding/setup_experiment.R"))

## Get SS
var_expt <- var_expt[1:2,]

```

## Load the old data

```{r load}
c1 <- readRDS("~/Documents_Local/git/diversity_envresp1/experiments/0_ss_finding/data/1c_9_strain_SS_data.RDS")
c6 <- readRDS("~/Documents_Local/git/diversity_envresp1/experiments/0_ss_finding/data/6c_9_strain_SS_data.RDS")
```

## The error

```{r error1}
c1 %>% mutate(stability_measures = list(get_stability_measures_new(ssfind_result)))
c6 %>% mutate(stability_measures = list(get_stability_measures_new(ssfind_result)))
```

## Further digging
```{r error2}
c1["ssfind_result"] %>% mutate(stability_measures = list(get_stability_measures_new(ssfind_result)))
c6["ssfind_result"] %>% mutate(stability_measures = list(get_stability_measures_new(ssfind_result)))
```

## Check the attributes
```{r}
attributes(c1["ssfind_result"])
attributes(c6["ssfind_result"])
```

The single core result is a `rowwise_df`, while the multicore is not.

## Fix
Add the `%>% rowwise()` to the function `run_temporal_ssfind_experiment()` inm the microxanox package and regenerate the data (in separate script)

## Load the new data

```{r loadNew}
c1_r <- readRDS("~/Documents_Local/git/diversity_envresp1/experiments/0_ss_finding/data/1c_rowwise_9_strain_SS_data.RDS")
c6_r <- readRDS("~/Documents_Local/git/diversity_envresp1/experiments/0_ss_finding/data/6c_rowwise_9_strain_SS_data.RDS")
```

## Now working

```{r Fixed}
c1_r %>% mutate(stability_measures = list(get_stability_measures_new(ssfind_result)))
c6_r %>% mutate(stability_measures = list(get_stability_measures_new(ssfind_result)))
```