### 2022-01-20 experiments/1_figures_main_ms/Understand_PB.Rmd
Moved figures in single Rmd document and added Illustrative document for PB (Figure 5)

---
title: "Understandin PB"
author: "Rainer M Krug"
date: "1/20/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = FALSE
)
library(microxanox)
library(tidyverse)
library(patchwork)
library(here)

colfunc_CB <- colorRampPalette(c("#024F17", "#B5FFC9"))
colfunc_SB <- colorRampPalette(c("#7D1402", "#FCBEB3"))
colfunc_PB <- colorRampPalette(c("#6E0172", "#F9AEFC"))

source(here("R/various_useful_functions.r"))

plot_fig3 <- function(
  ss_result
){
  temp <- ss_result$ssfind_result[[1]] %>%
    mutate(a = 10^a_O) %>%
    # arrange(a) %>%
    # mutate(direction = rep(c("up", "down"), nrow(ss_result)/2)) %>%
    #mutate(direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
    # select(-initial_N_CB, -a_O) %>%
    gather(species, quantity, 2:(ncol(.) - 2)) %>%
    mutate(
      var_type = ifelse(grepl("B_", species), "Organism", "Substrate"),
      functional_group = case_when(
        str_detect(species, "CB_") ~ "CB",
        str_detect(species, "SB_") ~ "SB",
        str_detect(species, "PB_") ~ "PB"
      ),
      log10_quantity = log10(quantity + 1)
    )

  num_strains <- temp %>%
    group_by(functional_group) %>%
    summarise(num = length(unique(species))) %>%
    na.omit()

  num_CB_strains <- num_strains$num[num_strains$functional_group == "CB"]
  num_SB_strains <- num_strains$num[num_strains$functional_group == "SB"]
  num_PB_strains <- num_strains$num[num_strains$functional_group == "PB"]

  line_width <- 2

  p1 <- temp %>%
    dplyr::filter(functional_group == "CB")  %>%
    mutate(species = factor(species, levels = unique(species))) %>%
    ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
    geom_path(lwd = line_width) +
    ##ylab("log10(quantity [cells])") +
    ylab("Log(Quantity)") +
    xlab("Oxygen diffusivity") +
    scale_colour_manual(values = colfunc_CB(num_CB_strains)) +
    guides(colour = guide_legend(ncol = 3)) +
    theme(legend.position="none") +
    ggtitle("(a) Cyanobacteria")

  p2 <- temp %>%
    dplyr::filter(functional_group == "SB")  %>%
    mutate(species = factor(species, levels = unique(species))) %>%
    ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
    geom_path(lwd = line_width) +
    ##ylab("log10(quantity [cells])") +
    ylab("Log(Quantity)") +
    xlab("Oxygen diffusivity") +
    scale_colour_manual(values = colfunc_SB(num_SB_strains)) +
    guides(colour = guide_legend(ncol = 3)) +
    theme(legend.position="none") +
    ggtitle("(b) Sulfate reducing bacteria")

  p3 <- temp %>%
    dplyr::filter(functional_group == "PB")  %>%
    mutate(species = factor(species, levels = unique(species))) %>%
    ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
    geom_path(lwd = line_width) +
    ##ylab("log10(quantity [cells])") +
    ylab("Log(Quantity)") +
    xlab("Oxygen diffusivity") +
    scale_colour_manual(values = colfunc_PB(num_PB_strains)) +
    guides(colour = guide_legend(ncol = 3)) +
    theme(legend.position="none") +
    ggtitle("(c) Phototrophic sulfur bacteria")

  p4 <- temp %>%
    dplyr::filter(var_type == "Substrate", species == "O") %>%
    ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
    geom_path(lwd = line_width) +
    ##ylab("log10(quantity [cells])") +
    ylab("Log(Quantity)") +
    xlab("Oxygen diffusivity") +
    theme(legend.position="none") +
    ggtitle("(d) Oxygen concentration")

  p5 <- temp %>%
    dplyr::filter(var_type == "Substrate", species == "SR") %>%
    ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
    geom_path(lwd = line_width, col = 6) +
    ##ylab("log10(quantity [cells])") +
    ylab("Log(Quantity)") +
    xlab("Oxygen diffusivity") +
    theme(legend.position="none") +
    ggtitle("(e) Sulfide concentration")
  patchwork <- p1 / p2 / p3 / p4 / p5
  patchwork
}
```

## Figure 5
Region of bistability by diversity and number of strains.

![alt text](figure_5_v2_allstrains.png)

## Figure 3
Stable state with diversity in all functional groups.

```{r fig3_prep}
ss_9s <- readRDS(here("experiments/0_ss_finding/temporal_method/data/9_strain_SS_data_1e6.RDS"))

max <- max(ss_9s$PB_var_gmax_s)
f <- 0.625
fv <- f*max

only_PB_variable <- ss_9s %>%
  filter(
    abs(CB_var_gmax_s) < 0.0001,
    abs(SB_var_gmax_s) < 0.0001
  )
only_PB_variable <- only_PB_variable[order(only_PB_variable$PB_var_gmax_s), ]

only_PB_variable$PB_var_gmax_s
```

The following are the same figure for different variabilities in PB. PB Ranges from `r min(ss_9s$PB_var_gmax_s)` to `r max(ss_9s$PB_var_gmax_s)`.

- PB_var_gmax_s: `r max(ss_9s$PB_var_gmax_s)`
- PB_var_gmax_s: `r max(ss_9s$PB_var_gmax_s)`
- PB_var_gmax_s: `r max(ss_9s$PB_var_gmax_s)`

### PB_var_gmax_s ascending
#### v = `r only_PB_variable$PB_var_gmax_s[1] / max(only_PB_variable$PB_var_gmax_s)`

```{r figure3_1, fig.height=7}
plot_fig3(only_PB_variable[1,])
```

#### v = `r only_PB_variable$PB_var_gmax_s[10] / max(only_PB_variable$PB_var_gmax_s)`

```{r figure3_10, fig.height=7}
plot_fig3(only_PB_variable[10,])
```

#### v = `r only_PB_variable$PB_var_gmax_s[12] / max(only_PB_variable$PB_var_gmax_s)`

```{r figure3_12, fig.height=7}
plot_fig3(only_PB_variable[12,])
```

#### v = `r only_PB_variable$PB_var_gmax_s[14] / max(only_PB_variable$PB_var_gmax_s)`

```{r figure3_14, fig.height=7}
plot_fig3(only_PB_variable[14,])
```

#### v = `r only_PB_variable$PB_var_gmax_s[16] / max(only_PB_variable$PB_var_gmax_s)`

```{r figure3_16, fig.height=7}
plot_fig3(only_PB_variable[16,])
```

#### v = `r only_PB_variable$PB_var_gmax_s[18] / max(only_PB_variable$PB_var_gmax_s)`

```{r figure3_18, fig.height=7}
plot_fig3(only_PB_variable[18,])
```

#### v = `r only_PB_variable$PB_var_gmax_s[20] / max(only_PB_variable$PB_var_gmax_s)`

```{r figure3_20, fig.height=7}
plot_fig3(only_PB_variable[20,])
```