### 2022-11-15 reports/animations/report.Rmd

---
title: "State change and diversity animations"
author: "Owen Petchey"
date: "6/1/2022"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(here)
```

# Cyanobacteria

## No diversity

```{r}
include_graphics(here("reports/animations/mp4s/CB_no_div.mp4"))
```

## Some diversity

```{r}
include_graphics(here("reports/animations/mp4s/CB_medium_div.mp4"))
```

# Sulfate reducing bacteria

## No diversity

```{r}
include_graphics(here("reports/animations/mp4s/SB_no_div.mp4"))
```

## Some diversity

```{r}
include_graphics(here("reports/animations/mp4s/SB_medium_div.mp4"))
```

# Phototrophic sulfur bacteria

## No diversity

```{r}
include_graphics(here("reports/animations/mp4s/PB_no_div.mp4"))
```

## Some diversity

```{r}
include_graphics(here("reports/animations/mp4s/PB_medium_div.mp4"))
```

# Oxygen concentration

## No diversity

```{r}
include_graphics(here("reports/animations/mp4s/O_no_div.mp4"))
```

## Some diversity

```{r}
include_graphics(here("reports/animations/mp4s/O_medium_div.mp4"))
```

# Reduced sulfur concentration

## No diversity

```{r}
include_graphics(here("reports/animations/mp4s/SR_no_div.mp4"))
```

## Some diversity

```{r}
include_graphics(here("reports/animations/mp4s/SR_medium_div.mp4"))
```

# Sulfate concentration

## No diversity

```{r}
include_graphics(here("reports/animations/mp4s/SO_no_div.mp4"))
```

## Some diversity

```{r}
include_graphics(here("reports/animations/mp4s/SO_medium_div.mp4"))
```

# Phospate concentration

## No diversity

```{r}
include_graphics(here("reports/animations/mp4s/P_no_div.mp4"))
```

## Some diversity

```{r}
include_graphics(here("reports/animations/mp4s/P_medium_div.mp4"))
```