### 2022-11-21 reports/stable_state_graphs/temporal/event_definition_2/3strain 1e6 wait_time/allssplots_strainvar_0.9.1.Rmd
mostly removing nonstandard stable state data

---
title: All stable state plots, with strain variation, 3 strains
author: "Owen L Petchey, Uriah Daugaard, Anubhav Gupta, Rainer M. Krug, Kimberley D. Lemmen, Sofia J. van Moorsel, Marcel Suleiman, Debra Zuppinger-Dingley, Romana Limberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    code_folding: hide
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(include = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE)
```

```{r echo = FALSE}
rm(list = ls())
library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("r", "various_useful_functions.R"))
# source(here("experiments", "1_figures_main_ms", "ms_figure_functions.R"))
source(here("R/ms_figure_functions.R"))

#print(fig_div_vs_o2diff_1strain_7row)

colfunc_CB <- colorRampPalette(c("#024F17", "#B5FFC9"))
colfunc_SB <- colorRampPalette(c("#7D1402", "#FCBEB3"))
colfunc_PB <- colorRampPalette(c("#6E0172", "#F9AEFC"))
zero <- 0
unity <- 1
```

```{r}
microxanox_version_here <- "0.9.1"
microxanox_version_path <- paste0("microxanox_v", microxanox_version_here)
if (packageVersion("microxanox") < package_version(microxanox_version_here)) {
  stop(paste0("microxanox version needs to be at least ", microxanox_version_here))
}
print(paste0("Version of microxanox: ", microxanox_version_here))
```

```{r}
wait_time <- "1e+06"
ss_9s <- readRDS(here::here("data",
                            microxanox_version_path, 
                            "temporal_method",
                            "event_definition_2",
                            paste0("ss_data_3strains_waittime", wait_time, "_event_definition_2.RDS")))

max_CB_var_gmax <- max(ss_9s$CB_var_gmax_s)
max_SB_var_gmax <- max(ss_9s$SB_var_gmax_s)
max_PB_var_gmax <- max(ss_9s$PB_var_gmax_s)

ss_9s <- ss_9s %>%
  mutate(standvar_CB_gmax = CB_var_gmax_s / max_CB_var_gmax,
         standvar_SB_gmax = SB_var_gmax_s / max_SB_var_gmax,
         standvar_PB_gmax = PB_var_gmax_s / max_PB_var_gmax)

```

# CB

```{r}
this_dat <- ss_9s %>%
  filter(
    SB_var_gmax_s == 0,
    PB_var_gmax_s == 0
  )
this_dat <- this_dat[order(this_dat$CB_var_gmax_s), ]

for(i in 1:20) {
  print(paste0("Standardised level of variation = ", round(this_dat$standvar_CB_gmax[i],2)))
  print(fig_state_vs_o2diff_sidebyside_alternative(this_dat[i,]))
}

#fig_state_vs_o2diff(only_SB_variable[13,])
```

# SB

```{r}
this_dat <- ss_9s %>%
  filter(
    CB_var_gmax_s == 0,
    PB_var_gmax_s == 0
  )
this_dat <- this_dat[order(this_dat$SB_var_gmax_s), ]

for(i in 1:20) {
  print(paste0("Standardised level of variation = ", round(this_dat$standvar_SB_gmax[i],2)))
  print(fig_state_vs_o2diff_sidebyside_alternative(this_dat[i,]))
}
#fig_state_vs_o2diff(only_SB_variable[13,])
```

# PB

```{r}
this_dat <- ss_9s %>%
  filter(
    CB_var_gmax_s == 0,
    SB_var_gmax_s == 0
  )
this_dat <- this_dat[order(this_dat$PB_var_gmax_s), ]

for(i in 1:20) {
  print(paste0("Standardised level of variation = ", round(this_dat$standvar_PB_gmax[i],2)))
  print(fig_state_vs_o2diff_sidebyside_alternative(this_dat[i,]))
}
#fig_state_vs_o2diff(only_SB_variable[13,])
```

# CB-SB

```{r}
this_dat <- ss_9s %>%
  filter(
    PB_var_gmax_s == 0,
    abs(standvar_CB_gmax - standvar_SB_gmax) < 0.00001
  )
this_dat <- this_dat[order(this_dat$CB_var_gmax_s), ]

for(i in 1:20) {
  print(paste0("Standardised level of variation = ", round(this_dat$standvar_CB_gmax[i],2)))
  print(fig_state_vs_o2diff_sidebyside_alternative(this_dat[i,]))
}
#fig_state_vs_o2diff(only_SB_variable[13,])
```

# CB-PB

```{r}
this_dat <- ss_9s %>%
  filter(
    SB_var_gmax_s == 0,
    abs(standvar_CB_gmax - standvar_PB_gmax) < 0.00001
  )
this_dat <- this_dat[order(this_dat$CB_var_gmax_s), ]

for(i in 1:20) {
  print(paste0("Standardised level of variation = ", round(this_dat$standvar_CB_gmax[i],2)))
  print(fig_state_vs_o2diff_sidebyside_alternative(this_dat[i,]))
}
#fig_state_vs_o2diff(only_SB_variable[13,])
```

# SB-PB

```{r}
this_dat <- ss_9s %>%
  filter(
    CB_var_gmax_s == 0,
    abs(standvar_SB_gmax - standvar_PB_gmax) < 0.00001
  )
this_dat <- this_dat[order(this_dat$SB_var_gmax_s), ]

for(i in 1:20) {
  print(paste0("Standardised level of variation = ", round(this_dat$standvar_SB_gmax[i],2)))
  print(fig_state_vs_o2diff_sidebyside_alternative(this_dat[i,]))
}
#fig_state_vs_o2diff(only_SB_variable[13,])
```

# CB-SB-PB

```{r}
this_dat <- ss_9s %>%
  filter(
    abs(standvar_CB_gmax - standvar_SB_gmax) < 0.00001,
    abs(standvar_SB_gmax - standvar_PB_gmax) < 0.00001
  )
this_dat <- this_dat[order(this_dat$SB_var_gmax_s), ]

for(i in 1:20) {
  print(paste0("Standardised level of variation = ", round(this_dat$standvar_CB_gmax[i],2)))
  print(fig_state_vs_o2diff_sidebyside_alternative(this_dat[i,]))
}
#fig_state_vs_o2diff(only_SB_variable[13,])
```