### 2022-11-25 reports/old/testing/testing_temporal_method.Rmd

---
title: "Testing temporal method"
author: "Owen Petchey"
date: "11/24/2021"
output: html_document
---

# Setup

```{r}
rm(list = ls())
library(here)
source(here("R/various_useful_functions.r"))

num_CB_strains <- 9
num_SB_strains <- 9
num_PB_strains <- 9

source(here("R/setup_experiment.R"))
```

## Version of `microxanox` package used: `r packageVersion("microxanox") `

## Display diversity

```{r}
display_diversity(
  400,
  var_expt = var_expt,
  num_CB_strains = num_CB_strains,
  num_SB_strains = num_SB_strains,
  num_PB_strains = num_PB_strains
)
```

## Test runs

```{r eval = FALSE}
## Here a test of one run
no_diverity_varexp_row <- 1
parameter$strain_parameter <- var_expt$pars[[no_diverity_varexp_row]]
parameter <- set_temporal_ssfind_initial_state(parameter,
                                               10^5,
                                               10^5,
                                               10^5)
no_div_res <- run_temporal_ssfind_method(parameter)
no_div_res
```

```{r eval = FALSE}
## here a test of an expt run
var_expt <- var_expt[1:2,]
expt_res <- run_temporal_ssfind_experiment(parameter,
                                           var_expt,
                                           total_initial_abundances = 10^5)
```

## Test plotting

```{r}
## test the plotting functions
plot_ss_result3(expt_res, 1)
plot_ss_result3(expt_res, 2)
plot_ss_result4(expt_res, 1, 2)

```

## Test get stability

```{r}
ss_object <- expt_res$ssfind_result[[1]]

get_stability_measures(expt_res$ssfind_result[[1]])

stab_data <-  expt_res %>%
  mutate(stability_measures = list(get_stability_measures(ssfind_result))) %>%
  unnest(cols = c(stability_measures))

```

## Test plot stability

Set up

```{r eval = plot_ss_results}
CB_vars <- unique(stab_data$CB_var_gmax_s)
SB_vars <- unique(stab_data$SB_var_gmax_s)

CB_stab_data <- stab_data %>%
  filter(SB_var_gmax_s == 0) %>%
  mutate(var_treat = "CB",
         var_gmax = CB_var_gmax_s)

SBPB_stab_data <- stab_data %>%
  filter(CB_var_gmax_s == 0) %>%
  mutate(var_treat = "SB-PB",
         var_gmax = SB_var_gmax_s)

for_join <- tibble(CB_var_gmax_s = sort(CB_vars),
                     SB_var_gmax_s = sort(SB_vars))
CBSBPB_stab_data <- stab_data %>%
  right_join(for_join) %>%
  mutate(var_treat = "CB-SB-PB",
         var_gmax = CB_var_gmax_s)

all_stab_results <- CB_stab_data %>%
  bind_rows(SBPB_stab_data) %>%
#  bind_rows(results3) %>%
 # bind_rows(results4) %>%
  bind_rows(CBSBPB_stab_data)

all_stab_results <- all_stab_results %>%
    mutate(var_treat = forcats::fct_relevel(var_treat, levels = c("CB", "SB-PB", "CB-SB-PB")))

```

### Log transformed

```{r eval = plot_ss_results}

all_stab_results %>%
  filter(Species == "O") %>%
  ggplot(aes(x = var_gmax, y = hyst_range_log, col=var_treat)) +
  geom_point() +
  xlab("Amount of trait variation\n[see text for units]") +
  ylab("Extent of bistability region\n[log10 oxygen diffusivity]") +
  labs(col = "Variation in\nonly these\nfunctional groups")
##ggsave("manuscript/figures/extent_of_bistab1.pdf", height = 4)

all_stab_results %>%
  #filter(var_treat == "CB") %>%
  filter(Species == "O") %>%
  ggplot(aes(x = var_gmax,
             ymin = hyst_min_log,
             ymax = hyst_max_log,
             fill=var_treat)) +
  geom_ribbon(alpha = 0.5) +
  facet_wrap( ~ var_treat, nrow = 3) +
  xlab("Amount of trait variation\n[see text for units]") +
  ylab("Oxygen diffusivity\n[log10 uM per hour]") +
  labs(fill = "Variation in\nonly these\nfunctional groups") +
  coord_flip() +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )

# ##ggsave("manuscript/figures/extent_of_bistab2.pdf", height = 4)
# stab_data %>%
#   filter(Species == "O",
#          sim_length == 1e6) %>%
#   ggplot(aes(x = CB_var_gmax_s, y = SB_var_gmax_s, fill = hyst_range_log, col = hyst_range_log)) +
#   geom_point(size = 3)
# 
# 
# ##ggsave("manuscript/figures/extent_of_bistab2.pdf", height = 4)
# 
# stab_data %>%
#   filter(Species == "O",
#          sim_length == 1e6) %>%
#   ggplot(aes(x = CB_var_gmax_s, y = hyst_range_log, col = as.factor(SB_var_gmax_s))) +
#   geom_line(size = 2)
```