### 2023-01-11 index.Rmd
new version figure 3, new figures in supplement

---
title: "Guide to repository of supplementary resources"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This is a guide to a repository containing various resources associated with the publication **Functional diversity can facilitate the collapse of an undesirable ecosystem state**.

[The repository on github, current version 0.9.1.](https://github.com/UZH-PEG/diversity_envresp1/releases/tag/0.9.1)

[The respository on Zenodo with DOI 10.5281/zenodo.7360272
](https://zenodo.org/record/7360272#.Y4CXmOzMJqs)

[The Supplementary Report (HTML page)](./reports/supplement/supplement_0.9.1.html)

[Instructions of how to reproduce the research (HTML page).](./docs/How_to_reproduce.html)

[Some animations of results (HTML page)](./reports/animations/report.html)