### 2023-02-01 index.Rmd
Update links in index.Rmd

---
title: "Guide to repository of supplementary resources"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This is a guide to a repository containing various resources associated with the publication **Functional diversity can facilitate the collapse of an undesirable ecosystem state**.

[The repository on github, current version 0.9.2.](https://github.com/UZH-PEG/diversity_envresp1/releases/tag/v0.9.2)

[The respository on Zenodo with the versioned DOI 10.5281/zenodo.7594024
](https://zenodo.org/record/7594024#.Y9pZaC8w0Q8)

[The Supplementary Report (HTML page)](./reports/supplement/supplement_0.9.1.html)

[Instructions of how to reproduce the research (HTML page).](./docs/How_to_reproduce.html)

[Some animations of results (HTML page)](./reports/animations/report.html)